---
title: "PML Prediction Assignment"
author: "richard-data"
date: "December 9, 2018"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load Datasets & Packages

Load in the testing and training dataset. Load in the following package.Take a look at the datasets.

```{r cars}
testdata<-read.csv("pml-testing.csv")
traindata<-read.csv("pml-training.csv")
library(caret)
library(rattle)
library(rpart)
library(rpart.plot)
library(randomForest)
library(repmis)
head(testdata,10)
head(traindata,10)
```

## Remove missing values and Split data
Remove the missing values from the datasets where more than 90% are null. In addition, remove the first seven columns with names and timestamps because they are not important.

Then, the training set will be subsetted (70/30) for training and validation. 

```{r}
coltoremove<-which(colSums(is.na(traindata) |traindata=="")>0.90*dim(traindata)[1])
training2<-traindata[,-coltoremove]
training3<-training2[,-c(1:7)]
testing3<-testdata[,-c(1:7)]

fortraining<-createDataPartition(training3$classe,p=0.70,list=FALSE)
train<-training3[fortraining,]
validate<-training3[-fortraining,]
```

##Classification Tree
```{r}
set.seed(12354)
trControl<-trainControl(method="cv",number=5)
model<-train(classe~.,data=train,method="rpart",trControl=trControl)
fancyRpartPlot(model$finalModel)
```

```{r}
trainpred<-predict(model,newdata=validate)
CONFUSION<-confusionMatrix(validate$classe,trainpred)
CONFUSION
```

The accuracy is 49%. Let's try another model. 

##Random Forest
```{r}
RANDOMFOREST<-train(classe~.,data=train,method="rf",trControl=trControl,verbose=FALSE)
print(RANDOMFOREST)
```
```{r}
trainpred2<-predict(RANDOMFOREST,newdata=validate)
confusion2<-confusionMatrix(validate$classe,trainpred2)
confusion2
```
The accuracy is very high and I'm willing to call it here.
Accuracy is 99.35% and the estimated out-of-sample error is 0.65%.

##PREDICT TEST DATA
```{r}
FINAL<-predict(RANDOMFOREST,newdata=testdata)
FINAL
```
